### 2013-08-20 docs/manuscript.tex
Added docs

%!TEX TS-program = pdflatex
\documentclass{emulateapj}

\usepackage{amsmath,amssymb}
\shorttitle{SCOPE}
\shortauthors{Casey et al}

\begin{document}

\title{SCOPE: Automatic Spectroscopic Analysis for The Poor Man}

\author{Andrew R. Casey\altaffilmark{1}}
\altaffiltext{1}{Research School of Astronomy \& Astrophysics, Australian National University, Mount Stromlo Observatory, via Cotter Rd, Weston, ACT 2611, Australia; \email{acasey@mso.anu.edu.au}}

\begin{abstract}

\end{abstract}

\section{Introduction}