### 2014-07-14 document/manuscript.tex
Writey writey hole

\documentclass{aastex}
\usepackage[utf8]{inputenc}

\newcommand{\sick}{\textit{sick}}
\newcommand{\article}{\textit{Article}}

\usepackage{amsmath}
\usepackage{bm}
\usepackage{natbib}
\usepackage{graphicx}

\begin{document}

\title{\sick, the spectroscopic inference crank}

\author{Andrew R. Casey\altaffilmark{1}}

\altaffiltext{1}{Institute of Astronomy, University of Cambridge, Madingley Road, Cambdridge, CB3 0HA, United Kingdom; \email{arc@ast.cam.ac.uk}}

% TODO:
% [ ] Create example code figures for the model
%  -- use pygmentize when necessary
% [X] Create some awesome GC spectral plots with Seaborn
%  -- [X] uncertainty in parameters as a function of S/N
%  -- [ ] Box plot distribution for the clusters
% [X] Run Sol at high-resolution?
%  -- [ ] plot the Sun
%  -- [ ] plot of random samples + optimization
%  -- [ ] plot triangle contours for the Sun

\begin{abstract}
In this \article{} I introduce \sick{}, the spectroscopic inference crank, an open source probabilistic code for inferring astrophysical parameters from spectra. \sick{} enables any user to easily construct an approximate \textit{generative model} for spectral data, allowing for precise inference of quantifiable properties for a host of astrophysical processes. Model fluxes (or intensities) are approximated by efficient multi-dimensional interpolation of pre-computed spectra. This allows any user to capitalise on the plethora of published synthetic and observed spectra available without extra effort. Additional phenomena that transform the data (e.g., redshift, continuum, smoothing) are incorporated as free parameters. Outlier pixels (e.g., cosmic rays or poorly modelled regimes) are treated with a mixture model, and a simplistic noise model is included to account for systematically underestimated variance. Combining these phenomena into a justified, quantitative model allows for precise inference with credible uncertainties. All aspects of the probabilistic model are introduced, and the implementation details are described. I demonstrate the utility of a generative model approach by precise stellar photospheric measurements from noisy (S/N $\sim{} 7$) high- and low-resolution spectra. \sick{} is very easy to use, well-tested, parallelised, and freely available online through GitHub under the MIT license. 
\end{abstract}

\section{Introduction}
Most of what we know about astrophysics has been interpreted from spectra. Given how informative spectroscopic data is to our understanding of astrophysics, it is not surprising that the last decade has seen a substantial increase of publicly accessible spectra. Large scale surveys have helped drive this trend, each releasing in excess of hundreds of thousands \citep[e.g.,][]{wigglez,boss,segue,rave} or even millions \citep{lamost,galah} of spectra.

The spectra are chosen using different source selections, and the data vary in wavelength coverage, resolution, and noise distributions. For these reasons many collaborations expend significant resources to produce bespoke analysis software to achieve their scientific objectives. Unfortunately this impedes scientific reproducibility, as many codes still remain closed-source more than a decade after the original article was published. Subsequently any comprehensive literature comparison effectively becomes impossible, as systematics can be difficult to accurately characterise without in-depth knowledge of the method or access to the software.

Broadly speaking, there are three types of methods employed for spectral analysis: measuring strength of spectral features, data-generating models or template-matching methods. Approaches that measure spectral features are inexpensive, but regularly encounter problems with blended (often hidden) lines or continuum placement. As such, some subjective interaction or tuning is often required. Data-generative methods compute model spectra at run-time, and whilst accurate, can be prohibitively expensive for large samples. For template-matching methods, synthetic spectra are only generated once to produce a grid of fluxes or intensities for each permutation of astrophysical quantities. Although there are differences between these methods, the preparatory steps are usually the same. Spectra are placed at rest-frame by calculating line-of-sight velocities, typically by cross-correlation, before being continuum-normalised to flux intensities.

Credible uncertainties can be difficult to discern from these approaches. The implied assumptions of a $\chi^2$ distribution are often broken, and uncertainties in the doppler-shift, smoothing, and continuum normalisation steps are almost always ignored. These effects result in ill-characterised uncertainties in astrophysical parameters. Alternatively the uncertainties are frequently, and often incorrectly, assumed to be dominated by systematic effects, and the magnitude of this effect is gauged by comparisons with a `trustworthy' literature sample. The simplest procedure is to then state all objects have roughly the same uncertainties in astrophysical parameters, since they are assumed to be dominated by systematics. This is an incorrect approach: there are few, if any, examples of homoscedastic datasets in astrophysics. The noise properties of each spectrum \textit{are} different, and the parameter uncertainties (random and systematic) will differ for every object. Consequently the uncertainties in astrophysical parameters by template-matching methods are generally found to either be incorrectly assumed, underestimated, or at least ill-characterised.

In addition to affecting the uncertainties, the effects of redshift, continuum normalisation and smoothing \textit{will} systematically bias the reported maximum-likelihood parameters. For example, there are a number of controversial examples within the literature where the subjective (human) decision of continuum placement has significantly altered the scientific conclusions \citep[e.g., see][where this issue is argued in great detail]{kerzendorf}. The implications of these phenomena \textit{must} be considered if we are to understand subtle astrophysical processes. Spectroscopy requires objectivity: one should endeavour to incorporate these phenomena as free parameters into a generative model and infer them simultaneously with our astrophysical parameters.

In this \article{} I present \sick{}: a well-tested, MIT-licensed probabilistic software package for inference from spectroscopic data. \sick{} employs an approximation to data-generating models: instead of attempting to model all manners of expensive astrophysical processes (e.g., supernova explosions, stars, and any other astrophysically interesting process) at runtime, it performs efficient multi-dimensional interpolation of model spectra from pre-computed grids. This approach is suitable for a plethora of different astrophysical processes, allowing any user to easily specify a model using any existing published spectral grid, and infer astrophysical properties from their data. The probabilistic model is described in Section \ref{sec:model}. A number of examples are included in Section \ref{sec:examples}, where I present accurate and precise objective photospheric measurements using noisy spectra. I conclude in Section \ref{sec:conclusions} with references to the online documentation and future applicability of the software.

\section{The Generative Model}

The primary advantage of \sick{} is the simultaneous inference of both astrophysical and other pertinent parameters that can alter the spectra. The generative model described here is agnostic as to \textit{what} the astrophysical parameters actually describe. Typical examples might be properties of supernova (e.g., explosion energies and luminosities), galaxy characteristics from integrated light, or mean plasma properties of a stellar photosphere. There are a plethora of synthetic spectral grids published for these types of applications \citep{who;who;who}. All of these models are fully-\sick{} compatible. However given the research background of the author, I will introduce the probabilistic model by considering the problem of inferring mean photospheric properties from a stellar spectrum.

The first step of the (approximate) generative model is to interpolate a model spectrum. \sick{} employs the Qhull algorithm \citep{qhull} to linearly interpolate in multiple dimensions. Irregularly-spaced grids are perfectly acceptable; Qhull does not require the grid points to be rectangular or regularly spaced. Large grids of model spectra can be cached for computational efficiency (and this is recommended), allowing for the total size of the model grid to far exceed the available random access memory.\footnote{In practice this is performed through memory-mapping.} I introduce an interpolation function $S(\bm{\psi})$ that produces a model spectrum with $J$ points $\{\lambda_j,I_j\}_{0}^{J}$ at any evaluable point $\bm{\psi} \equiv (T_{\rm eff},\log{g},{\rm [M/H]},[\alpha/{\rm Fe}])$ within the grid boundaries.

\begin{equation}
S(\bm{\psi}) \rightarrow \{\lambda_j,I_j\}
\end{equation}

The model spectrum $S(\bm{\psi})$ should always be of higher spectral resolution than the data. Therefore it is necessary to smooth (or broaden) $\{\lambda_j,I_j\}$ with a Gaussian filter of standard deviation $\sigma_{s}$ such that it matches the spectral resolution of the data. For some applications it is tempting to perform this step \textit{a priori} from the instrument spectral resolution. However it is not necessarily true that the quoted instrumental resolution will perfectly match the data. In truth the data have a \textit{distribution} of spectral resolutions. Even if the spectral resolution is accurately known with low variance, fixing this value will result in under-estimated uncertainties $\sigma(\bm{\psi})$ and likely bias the posterior distribution distributions for $\bm{\psi}$. Employing the free parameter $\sigma_{s}$ allows for the identification of additional instrumental or atmospheric broadening. For stellar applications, the free parameter $\sigma_{s}$ also aids in the identification of unresolved binary companions.

The doppler shift of the source is modelled by transforming the model spectrum (in uniformly sampled $\log\lambda$ space). For computational efficiency \sick{} solves for redshift $z$, and returns posterior distributions in the user's preferred units (e.g., km s$^{-1}$).

\begin{equation}
\lambda_{s} = \lambda_{j}(1 + z)
\end{equation}

After the model spectrum is smoothed and shifted, model intensities are required at the wavelengths $\lambda_i$ of the observed pixels. The complete function to yield model intensities $M_i$ at wavelengths $\lambda_i$ is therefore a function of the parameters of interest $\bm{\psi}$, redshift $z$, smoothing kernel $\sigma_{s}$, and the wavelengths of the observed pixels $\lambda_i$:

\begin{equation}
 M_{i} \equiv S\left(\bm{\psi},z,\sigma_{s},\lambda_i\right)
\end{equation}

In principle there is no reason to believe that the doppler-shifted, smoothed model intensities $M_i$ should match the data at all. The counts and shape of observed fluxes are a function of source magnitude, exposure time, instrument sensitivities, atmospheric conditions, and a host of unaddressed effects. In contrast the model spectra are calculated either as intensities (e.g., $I_j$) or calibrated fluxes to an empirical system. Even for the true values of $\bm{\psi}$, a function is required to normalise\footnote{The normalisation process is frequently abused by stellar spectroscopists in the literature. Wherever possible, data should not be transformed. One should seek to fit a model to the data, not the other way around.} the model to the data. As such I transform the \textit{model} by some function $C$ to fit the data. Although the function $C$ incorporates a number of effects (e.g., source blackbody temperature, dust, instrument sensitivities), they are phenomena that typically cannot be separated without additional information, and here I only care about their combined effect. I wish to ensure that the overall shape of the data is accounted for, which is usually achievable with a low-order polynomial

\begin{equation}
C_i = c_{0}\lambda_{i}^{m-1} + c_{1}\lambda_{i}^{m-2} + \dots + c_{m}
\end{equation}

\noindent{}with $m$ free coefficients $c_{0\dots{}m}$, where $m$ is specified by the user. Other continuum functions are available in \sick{}, and the user is encouraged to experiment. This allows us to express the \textit{expected flux} $E_i$ at a given observed pixel with wavelength $\lambda_i$ as:

\begin{equation}
E_i = S_{i}\times{}C_i
\end{equation}

In many astrophysical cases the flux uncertainties $\sigma_{i}$ for a given pixel are not well-characterised. This is often due to unpropagated uncertainties during data reduction. Since the observed flux counts are usually Poisson distributed, in many astrophysical scenarios the pixel flux uncertainty can be made estimated as $\sigma_i \sim 1/\sqrt{F_{i}}$. Knowing that this approximation may under-estimate the noise, I include an additional parameter to account for the possibility that the variance in all pixels is underestimated by some fractional amount $f$ (Eq. \ref{eq:noise-model}). While this crude noise model is probably unrepresentative of the data in a large number of cases, including \textit{any} noise model is preferable to none.

\begin{equation}
\sigma_{t_i}^2 = \sigma_{i}^2(1 + f)
\label{eq:noise-model}
\end{equation}

We now have a \textit{generative model} for the data. The frequency (or probability) distribution ${p\left(F_i|\lambda_i,\sigma_i,\bm{\psi},z,\sigma_s,\{c\}_{0}^{m}\right)}$ for $F_i$ is given by

\begin{equation}
p\left(F_i|\lambda_i,\sigma_{i},\bm{\psi},z,\sigma_{s},\{c\}_{0}^{m},f\right) = 
 \frac{1}{\sqrt{2\pi\sigma_{t_i}^2}}\exp{\left(-\frac{\left[F_i - E_i\right]^2}{2\sigma_{t_i}^2}\right)}
 \label{eq:p_model}
\end{equation}

\noindent{}and with the implied assumption that the data are independently drawn, the likelihood $\mathcal{L}$ is calculable by the product of individual probabilities:

\begin{equation}
\mathcal{L} = \prod_{i=1}^{N}\,p\left(F_i|\lambda_i,\sigma_{i},\bm{\psi},z,\sigma_{s},\{c\}_{0}^{m},f\right)
\end{equation}

From the original astrophysical parameters $\bm{\psi}$ that we care about, we now have an additional $3 + m$ parameters to consider. The situation seemingly becomes more complex when we consider separately observed channels (also colloquially expressed as beams or apertures). Many spectrographs provide small portions of spectra (a channel) separated by some large gap. The data for each channel are generally captured by different CCDs and are therefore reduced separately.

The instrumental parameters $\sigma_{s}$, $\{c_k\}_{k=0}^{m}$, and $f$ are likely to be different for each channel. Although redshift $z$ is an astrophysical effect and should not differ between channels, this may not be true if each channel has been wavelength-calibrated differently. It is therefore prudent to introduce separate parameters $z$, $\sigma_{s}$, $\{c_k\}_{k=0}^{m}$, $f$ for \textit{each} of the $N_{chan}$ observed channels. This scales the total dimensionality of the model as $3\times{}N_{chan} + \sum_{k=0}^{N_{chan}}m_{k}$ in addition to $\bm{\psi}$. The inclusion of all of these parameters is not mandatory: each \sick{} model can be adjusted to include or ignore any combination of phenomena.

Finally, I consider the handling of outliers in the data. These may be in the form of cosmic ray spikes, improper calibration of the data (e.g., skylines) or poorly modelled spectral regions. Treatment of these artefacts is achieved with a Gaussian mixture model, a combination of two models. This requires the inclusion of two (final) additional parameters: $P_o$ and $V_o$. In the mixture model, the data are fit by the sum of amplitudes ($1 - P_o$ and $P_o$, respectively) of two distributions: the expected fluxes $E_i$, and a normal distribution centered along the continuum function $C_i$ with variance $\sigma_{t_i}^2 + V_{o}$. The prior distribution function $p\left(V_o\right)$ requires $V_{o}$ to always be positive (Eq. \ref{eq:default_priors}), and as such the outlier distribution will \textit{always} have a larger variance. Because distributions of smaller variance are more informative, a fit to the expected fluxes $E_{i}$ is generally preferred wherever possible. For brevity I define $\bm{\kappa} \equiv (\bm{\psi},\{z,\sigma_s,\{c_k\}_{k=0}^{m},f\}_{0}^{N_{c}})$, and the likelihood for the mixture model is given by

 \begin{equation}
\mathcal{L} = \prod_{i=1}^{N}\,\left[\left(1 - P_{o}\right)\times{}p_{model}\left(F_i|\lambda_i,\sigma_{i},\bm{\kappa}\right) + P_{o}\times{}p_{outlier}\left(F_i|\lambda_i,\sigma_i,\bm{\kappa},V_{o},P_o\right)\right]
\end{equation}

\noindent{}where $p_{model}$ refers to $p$ in Eq. \ref{eq:p_model} and

\begin{equation}
p_{outlier}\left(F_i|\lambda_i,\sigma_i,\bm{\kappa},V_{o},P_o\right) = \frac{1}{\sqrt{2\pi\left(\sigma_{t_i}^2 + V_{o}^2\right)}} \exp\left(-\frac{[F_i - C_i]^2}{2\left[\sigma_{t_i}^2 + V_{o}^2\right]}\right)
\end{equation}

\noindent{}such that the likelihood $\mathcal{L}$ becomes:

\begin{equation}
\mathcal{L} = \prod_{i=1}^{N} \left[ \frac{1-P_o}{\sqrt{2\pi\sigma_{t_i}^2}}\,\exp\,\left(-\frac{[F_i - E_i]^2}{2\sigma_{t_i}^{2}}\right) + \frac{P_o}{\sqrt{2\pi\left[\sigma_{t_i}^2 + V_o\right]}}\,\exp\,\left(-\frac{[F_i - C_i]^2}{2\left[\sigma_{t_i}^{2} + V_o\right]}\right)\right]
\label{eq:full_likelihood}
\end{equation}

I define the full parameter space with $\bm{\theta} \equiv \left(\bm{\psi},\{z,\sigma_s,\{c_{b_k}\}_{k=0}^{m},f\}_{b=0}^{N_{c}},V_o,P_o\right)$. From Bayes theorem we obtain the posterior probability distribution for $\bm{\theta}$ (up to a constant) as
\begin{eqnarray}
\mathcal{P} & \propto & likelihood \times prior \nonumber \\
p(\bm{\theta}|\{F_i\}_{i=1}^{N}) & \propto & p(\{F_i\}_{i=1}^{N}|\bm{\theta})\,\times\,p(\bm{\theta})
\label{eq:probability}
\end{eqnarray}

\noindent{}where $p(\bm{\theta}|\{F_i\}_{i=1}^{N})$ is the probability $\mathcal{P}$ of $\bm{\theta}$ given the data (and the model: see Section \ref{sec:sun}), $p(\{F_i\}_{i=1}^{N}|\bm{\theta})$ is our previously defined likelihood function $\mathcal{L}$ (Eq. \ref{eq:full_likelihood}), and $p(\bm{\theta})$ is the prior probability distribution. Priors are discussed in more detail in Section \ref{sec:priors}.

%

% Log probability, expand function
%\begin{eqnarray}
%\log({\mathcal{P}}) & \propto & \log{likelihood} + \log{prior} \nonumber \\
%\log(\mathcal{P}) & \propto & \prod_{i=1}^{N} \left[ \frac{1-P_b}{\sqrt{2\pi\sigma_{i}^2}}\,\exp\,\left(-\frac{[F_i - E_i]^2}{2\sigma_{i}^{2}}\right) + \frac{P_b}{\sqrt{2\pi\left[\sigma_i^2 + V_o\right]}}\,\exp\,\left(-\frac{[F_i - C_i]^2}{2\left[\sigma_i^{2} + V_o\right]}\right)\right] \nonumber \\
%& & \dots +  \ln(p(\bm{\theta}))
%\end{eqnarray}

With all of the combined effects, there are 16 free parameters including $\bm{\psi}$, our astrophysical quantities of interest. For the examples presented in Section \ref{sec:examples}, this requires interpolation between $\sim{}92\times10^9$ pixels in four dimensions. While the description might appear daunting, the problem is tractable, numerically efficient, and easy to configure. By default \sick{} numerically solves this problem in three sequential steps, which are described in the following sections: scattering, optimisation, and Monte-Carlo Markov Chain (MCMC) sampling.

\subsection{Initial Scattering}
\label{sec:scattering}

We seek to maximise $\mathcal{P}$ (or in practice, $\log\mathcal{P}$) and calculate the posterior probability distributions for $\bm{\theta}$ given the data. In the first step we calculate the log-probability $\log{(\mathcal{P})}$ for $N_{sample}$ randomly drawn points from all over the astrophysical parameter space $\bm{\psi}$. Since \sick{} allows for arbitrarily large parameter spaces in $N_{D}$ dimensions, initially sampling the parameter space $N_{sample}$ times provides a coarse representation of probability. If the probability distribution function were smoothly distributed across all $\bm{\psi}$, or the optimisation step was sufficiently robust for all potential applications, then in principle \textit{any} single point would be an adequate starting guess. However there is no way of knowing \textit{a priori} how smooth the probability landscape is for a given problem.

The $N_{samples}$ points are uniformly drawn in $\bm{\psi}$ by default, unless priors are specified (see Section \ref{sec:priors}). A model spectrum $S(\bm{\psi})$ is interpolated for each point, and is used to estimate the redshift and continuum parameters. After smoothing the model fluxes $I_j$ by $\sigma_s$ (as drawn from an explicit prior, or estimated by $\left|\mathcal{N}\left(0, 1\right)\right|$) the smoothed spectrum is cross-correlated with the data to yield a redshift $z$. Continuum parameters $\{c_k\}_{k=0}^{m}$ are similarly estimated by fitting a function (e.g., a polynomial) to the data divided by the smoothed flux at each $\lambda_i$. Estimating continuum and redshift provides us with a reasonable guess of the probability for any randomly scattered point $\bm{\psi_i}$. If outlier modelling is included in the model then $P_o$ is distributed as $\mathcal{U}\left(0, 1\right)$ by default, and $V_o$ is estimated by $\mathcal{N}\left(\widetilde{F_i}, \frac{1}{2}\widetilde{F_i}\right)$ and $\mathcal{N}\left(\mbox{Var}\left(F_i\right),\frac{1}{2}\mbox{Var}\left(F_i\right)\right)$ respectively. This procedure is \textit{only} employed for the initial random scattering stage: it is \textit{not} applicable for MCMC sampling.

\subsection{Optimisation}
\label{sec:optimise}

After the random scattering step, the most probable $\bm{\theta}$ point is used as an initial guess for numerical optimisation. A number of suitable minimisation algorithms are available in \sick{} through the SciPy optimization \cite{scipy,scipy-optimize} module. The Nelder-Mead algorithm \citep{nelder-mead} is employed by default, and the full suite of optimisation algorithms available is listed in the online documentation. Unlike other optimisation techniques, the Nelder-Mead algorithm does not approximate first- or second-order derivatives, and can consequently be less efficient than other approaches. However it is robust in high-dimensional space, even in the presence of substantial noise, and has been successfully employed in a wide range of engineering and scientific problems. The reader is encouraged to experiment with other optimisation algorithms if the Nelder-Mead algorithm proves unsuitable or takes an untenable amount of time. Since these optimisation algorithms are minimisation techniques and we seek to maximise the log-probability $\log{\left(\mathcal{P}\right)}$, we numerically optimise the parameters $\bm{\theta}$ by minimising the negative log-probability $-\log{\left(\mathcal{P}\right)}$.

\subsection{Monte-Carlo Markov Chain Sampling}
\label{sec:mcmc}

The random scattering and numerical optimisation steps efficiently provide an accurate estimate of the optimal parameters $\bm{\theta}$. Once the optimisation step is complete \sick{} employs the affine-invariant Metropolis-Hastings MCMC ensemble sampler proposed by \citet{goodman;weare}, and implemented by \citet{emcee}. If the optimisation steps is not employed, initial points for the MCMC walkers are drawn in the same way as described in the scattering process. When the optimisation step is used, the initial points are taken from a small multi-dimensional ball around the optimised parameters $\bm{\theta_{opt}}$.

% [TODO] "Experiment" with proposal widths.

Priors represent our initial knowledge about a particular parameter before looking at the data, and are necessary for any Bayesian analysis. A number of different prior distributions can be specified by the user in the \sick{} model configuration file. When no prior is specified, the following uninformative prior distributions are assumed (for all channels, where appropriate):

\begin{eqnarray}
p\left(\bm{\psi}_{dim}\right) &=& \mathcal{U}\left(\min\left[\bm{\psi}_{dim}\right], \max\left[\bm{\psi}_{dim}\right]\right) \\
p\left(P_o\right) &=& \mathcal{U}\left(0, 1\right) \\
p\left(\log{f}\right) &=& \mathcal{U}\left(-10, 1\right) \\
p\left(V_o,\sigma_s\right) &=& \left\{
\begin{array}{c l}      
    1\,, &\mbox{for values greater than zero}\\
    0\,, &\mbox{otherwise}
\end{array}\right. \\
p\left(z,\{c_k\}_{k=0}^{m}\right) &=& 1 \\
\label{eq:default_priors}
\end{eqnarray}

A consequence of allowing irregular model grids is that a model spectrum cannot be interpolated for some values of $\bm{\psi}$, even if they fall within $\left(\min\left[\bm{\psi}\right], \max\left[\bm{\psi}\right]\right)$. In these cases $p\left(\bm{\psi}\right) = 0$ and thus $\log\left(\mathcal{P}\right) = -\infty$.

Quantitatively ensuring numerical convergence for MCMC analyses remains an unsolved problem. There are a number of excellent resources on MCMC sampling, which outline this issue in greater detail. The mean acceptance fraction, auto-correlation times, and values of the parameter chains themselves all provide reasonable proxy indicators of convergence. The user is encouraged to inspect these metrics and decide upon some heuristic for when convergence has unambiguously been achieved. The final state of every MCMC simulation is automatically saved by \sick{}. This allows users to `resume' their analysis from the most recent state if they believe the system has not converged.

\subsection{Self-consistent inference test}

In this test I have interpolated a noise-free synthetic spectrum from the blue (475\,nm to 685\,nm) channel of the AMBRE synthetic library \citep{ambre} which will act as a faux observation. The interpolated spectrum has had a number of transformations applied: a second-order polynomial that enters multiplicatively to represent the continuum, redshift, and convolved to a spectral resolution of $\mathcal{R} \sim 10000$. The faux data are resampled with a uniform spacing of $\sim{}0.8$\,\AA{}, effectively discarding $99.5\%$ of the original pixels, which were sampled at $\sim4\times10^{-4}$\,nm. White noise has been added to replicate a S/N ratio of $\sim 7$ px$^{-1}$. I also assumed that the observed variance is systematically underestimated by 10 per cent: conceptually the faux data is $\sim{}$10 per cent noisier than what an observer would estimate from the flux counts.

% The AMBRE synthetic spectra have been calculated using 1D MARCS model atmospheres and synthesised using Turbospectrum \citep{turbospectrum} under the assumption of local thermal equilibrium (LTE). There are 28,344 permutations of effective temperature $T_{\rm eff}$, surface gravity $\log{g}$, overall metallicity $[{\rm M/H}]$ and mean $\alpha$-element enhancement $[\alpha/{\rm Fe}]$. The typical step size is 250 K in temperature, 0.25 dex in $\log{g}$, 0.25 dex in [Fe/H] and 0.10 dex in [$\alpha$/Fe]. Each grid point has synthesised continuum-normalised intensities between [0, 1]  for two spectral channels: one from 475\,nm to 685\,nm, and a redder channel covering 845\,nm to 895\,nm. Both channels are sampled at approximately $\sim4\times10^{-4}$\,nm, yielding more than $92\times10^9$ pixels.

Default priors and configurations were employed for this example. 200 walkers were used to explore the parameter space for 1250 steps as burn-in, and another 250 steps were used to sample the posterior. The chains for each parameter are shown in Figure \ref{fig:chains}. At least by eye, the simulation appears to have converged after 300 steps. The marginalised posterior distributions are shown in Figure \ref{fig:corner-inference}, with the accepted truth values marked. The true and inferred values for all parameters is tabulated in Table \ref{tab:inference-test}.

Even in the presence of substantial noise, precise inference of stellar parameters can be made. The corner plots demonstrate a strong covariance between the continuum coefficients $c_0$, $c_1$ and $c_2$. This is unsurprising given the shape of the spectrum (Figure \ref{fig:spectrum-inference}), which could reasonably be fit with a linear continuum (even though a second-order polynomial was applied).

% Discuss precise infererence of ln(f_blue) -- not representative of real world data.

% Confident that the machinery works, allows extension to include other noise models and telluric absorption or resolve of binary systems?

\begin{figure}
\label{fig:chains}
\includegraphics[height=\textheight]{chains.pdf}
\caption{Points sampled by the 200 walkers in the self-consistent inference test. The true values are marked in blue. I discard the first 1250 steps as the burn-in period, and sample the posterior distribution function for 250 steps.}
\end{figure}

\begin{figure}
\label{fig:corner-inference}
\includegraphics[width=\textwidth]{corner.pdf}
\caption{Marginalised posterior distributions for all parameters $\bm{\theta}$ from a faux observation with spectral resolution $\mathcal{R} \sim 10,000$ and S/N ratio $\sim{}7$. True values are marked in blue. This figure demonstrates that precise inference of stellar parameters can be made with high-resolution spectra, even in the presence of substantial noise.}
\end{figure}

\begin{table}
\caption{True and inferred model parameters for the self-consistent inference test}
\begin{tabular}{llrr}
\hline
\hline
Parameter & Description & Truth & Inferred \\
\hline
$T_{\rm eff}$ & Effective photospheric temperature (K) & 5454 & 5483$_{-137}^{+125}$ \\
$\log{}g$ & Surface gravity & 4.124 & 4.084$_{-0.274}^{+0.262}$ \\
${\rm [Fe/H]}$	& Metallicity & --0.514 & $-0.512_{-0.160}^{+0.156}$ \\
$[\alpha/{\rm Fe}]$ & $\alpha$-element abundance & $+0.02$ & $+0.05_{-0.12}^{+0.13}$ \\
$v$		& Velocity (km s$^{-1}$) & 13.0 & $9.5_{-5.5}^{+5.3}$ \\
$\sigma_s$ & Gaussian smoothing sigma (\AA{}) & 0.581 & 0.518$_{-0.34}^{+0.39}$ \\
$\log{f}$ & Logarithm of fractionally underestimated variance & $-2.30$ & $-2.28_{-0.02}^{+0.02}$ \\
$c_{0}$ & Continuum polynomial coefficient ($\times10^{-3}$) & 1.23 & $1.26_{-0.32}^{+0.32}$ \\
$c_{1}$ & Continuum polynomial coefficient & $-0.593$ & $-1.001_{-3.634}^{+3.681}$ \\
$c_{2}$ & Continuum polynomial coefficient & $-0.569$ & $894_{-10335}^{+10367}$ \\
\hline
\end{tabular}
\end{table}

\begin{figure}
\label{fig:spectrum-inference}
\includegraphics[width=\textwidth]{spectrum.pdf}
\caption{A faux observed spectrum (black) of $\mathcal{R} \sim 10000$ and S/N ratio of $\sim7$ with an underestimated variance by $\sim$10\%. The maximum likelihood model spectrum is shown in red. }
\end{figure}

\section{Utility}
Here I include two realistic applications to demonstrate the capabilities and ease-of-use of the software. Recall that \sick{} is ambivalent about \textit{what} the astrophysical parameters are describing. However given the research background of the author, these examples will continue to focus on stellar applications.

\subsection{Sol}
A high-resolution ($\mathcal{R} \sim$ 20000) twilight spectrum was obtained from the GIRAFFE/FLAMES archive\footnote{https://www.eso.org/observing/dfo/quality/GIRAFFE/pipeline/solar.html}. A 25 second exposure using the H875.7 setup yields spectra from 848.2\,nm to 899.2\,nm. This setup was chosen as it most closely represents the wavelength region of the Gaia (847-874\,nm) survey. Data points redder than 874\,nm have been discarded to more closely match the Gaia spectral coverage. The data $D_n$ are represented by wavelengths\footnote{Spectral data $D_n$ can be represented as frequencies or any other unit equivalency.} and fluxes at $N$ pixels $\{\lambda_i,F_i\}_{0}^{N}$, and the noise in $\{F_i\}$ is Poisson-distributed with variance $\{\sigma_{i}^{2}\}_{0}^{N}$.

% DISCUSS DIFFERENCE IN TRUTH-INFERRED VALUES

% DISCUSS THE INCLUSION OF TELLURIC MODELLING

\subsection{Globular and Open Clusters}
Globular and open clusters are excellent candidates for testing this probabilistic framework over a large range of model parameters. Here I consider a realistic scenario of low-resolution, noisy spectra of cluster candidates observed using the AAOmega spectrograph on the Anglo-Australian Telescope in Coonabarabran, Australia. The 580V and 1700D gratings were employed, yielding one channel between X to Y nm with a spectral resolution Z, and another from X to Y nm with spectral resolution Z.

Although these data are generally considered to be of low spectral resolution, I show that astrophysical parameters can be accurately inferred with high precision. The AMBRE synthetic grid has also been employed for this example. Each observed channel was treated identically: individual doppler shifts were permitted in each channel, and the continuum was represented as a second-order polynomial. Outlier modelling was included. The central $\pm$0.1 nm cores of the Ca II near-infrared triplet lines were masked (specifically rest wavelengths 849.7-849.9 nm, 854.1-854.3, and 866.1-866.3 nm), as the cores are strongly affected by non-LTE effects that are not treated in the 1D LTE model atmospheres used to generate the AMBRE spectra.

The parameter space was randomly sampled 1000 times before numerical optimisation was performed using the Nelder-Mead algorithm. 200 walkers explored the parameter space for a burn-in period of 450 steps, before sampling the posterior another 50 times. Plots showing auto-correlation times, chain samples, and mean acceptance fractions demonstrated that convergence was unambiguously achieved for all examined sources.

Radial velocities were used to identify and discard non-cluster members. The metallicity distribution of the remaining cluster stars are shown in Figure X, where a comparison to the literature is made. The \citet{harris} catalogue (accessed June 2014) has been used as a literature reference for all globular clusters. For the open cluster Mel 66, the mean metallicity of three high-quality studies devoted to this system were used: $\langle[\mbox{Fe/H}]\rangle = -0.44$ as averaged from $-0.43$ \citep{who}, $-0.44$ \citep{who}, and $-0.45$ \citep{who}.

% Figure of metallicity comparison

The agreement between the literature and our measurements is unprecedented for low-resolution spectra. No systematic trend is present and a negligible mean metallicity difference of $-0.001$ dex (0.2\%) is observed over three orders of magnitude. The agreement would be remarkable for high-resolution spectra with exquisite S/N ratios. I emphasize that these results were obtained without performing any `calibration' or ad-hoc posterior corrections, which are usually entirely empirical and not motivated by an astrophysical understanding, I obtain superior precision and accuracy than alternative methods presented in the literature.

Individual uncertainties in stellar parameters are also extremely reasonable, particularly given the low spectral resolution and $S/N$ of the data. The uncertainties in stellar parameters are shown in Figure \ref{fig:gc-snr} as a function of $S/N$. Uncertainties of $\sim$75 K in $T_{\rm eff}$, $\sim0.2$ in $\log{g}$, $\sim$0.1 dex in [Fe/H] and $\sim0.05$ dex in [$\alpha$/Fe] are achievable even in the presence of substantial noise (e.g., $S/N \sim 15$). The \sick{} configuration file used to produce these results is shown in Figure Xa, and can be readily applied to large numbers of similar low-resolution spectra (e.g., RAVE, SEGUE, LAMOST, Gaia, or archival spectra from any number of telescopes).

%\begin{figure}
%\label{fig:gc-snr}
%\includegraphics[width=\textwidth]{figures/sick-gc-snr.pdf}
%\caption{Maximum absolute stellar parameter uncertainties from low-resolution spectra of globular and open cluster candidate stars as a function of signal-to-noise $S/N$ ratio. $S/N$ is calculated per pixel from the red channel.}
%\end{figure}

% Figure Xa shows a code block where (on left):
% show the model
% and on right:
% pip install sick
% sick get segue-model
% ls *
% sick solve segue-model.yaml segue-spectra-A.fits segue-spectra-B.fits

\section{Conclusion}

A probabilistic code has been introduced to approximately forward model spectroscopic data, and thereby infer astrophysical and nuisance parameters. The generative model approach described here has a number of advantages over previously published techniques. Preparatory and subjective decisions (e.g., redshift and placement of continuum) are objectively treated within a scalar-justified mathematical model. This allows us to credibly characterise the parameter posterior distributions and understand how they affect our astrophysical measurements. Until now, most published techniques treat these processes separately, thereby systematically biasing and generally under-estimating the parameter uncertainties.

With applications to both high- and low-resolution stellar spectra, I have demonstrated that the simultaneous incorporation of these effects leads to remarkable improvements in both accuracy and precision. While the examples have focussed on stellar spectra, the code can be equally used for any kind of quantifiable astrophysical process. The code is MIT-licensed (freely distributable, open source) and has an extensive automatic testing suite. Complete documentation is available online\footnote{github.com/andycasey/sick/wiki}, and is complemented with the files necessary to reproduce the examples outlined in this \article{}. The source code is distributed using git\footnote{git-scm.com/downloads}, and hosted online at GitHub\footnote{github.com/andycasey/sick}.

Great care has been spent to ensure the code is easy-to-use, and allow users to obtain precise inference with little effort. I therefore strongly encourage the use of this software for existing and future spectral datasets, as the need for precise and reproducible spectroscopic analysis has arrived. With the wealth of spectral data becoming available in the coming decade, astronomers must begin to adopt more objective, probabilistic modelling for their data. Subtle astrophysical processes can only be discovered and understood with the proper characterisation of uncertainties afforded by forward modelling.

\acknowledgements
The author is pleased to thank Sergey Koposov for constructive discussions on this work, and Louise Howes for extensive testing of the code. This research has been funded by European Research Council grant 320360: The Gaia-ESO Milky Way Survey.

\end{document}