### 2014-08-09 document/manuscript.tex
Updated manuscript

\documentclass{aastex}
\usepackage[utf8]{inputenc}

\newcommand{\sick}{\textit{sick}}
\newcommand{\article}{\textit{Article}}

\usepackage{amsmath}
\usepackage{bm}
\usepackage{natbib}
\usepackage{graphicx}

\begin{document}

\title{\sick, the spectroscopic inference crank}

\author{Andrew R. Casey\altaffilmark{1}}

\altaffiltext{1}{Institute of Astronomy, University of Cambridge, Madingley Road, Cambdridge, CB3 0HA, United Kingdom; \email{arc@ast.cam.ac.uk}}

\begin{abstract}
In this \article{}  I  introduce \sick{}, the spectroscopic inference crank, an open source probabilistic code for inferring astrophysical parameters from spectra. \sick{} enables any user to easily construct an approximate \textit{generative model} for spectral data, allowing for precise inference for a host of astrophysical quantities. Model fluxes (or intensities) are approximated by efficient multi-dimensional interpolation of pre-computed spectra. This allows any user to capitalise on the plethora of published synthetic and observed spectral libraries with minimal effort. Additional phenomena that transform the data (e.g., redshift, continuum, smoothing) are incorporated as free parameters. Outlier pixels (e.g., cosmic rays or poorly modelled regimes) can be treated with a mixture model, and a simplistic noise model is included to account for systematically underestimated variance. Combining these phenomena into a scalar-justified, quantitative model allows for precise inference with credible uncertainties. Commonly employed features are introduced, and the implementation details are described. I demonstrate the utility of a generative model approach with accurate and precise stellar photospheric measurements from noisy (e.g., S/N $\sim{} 7$) high- and low-resolution spectra. \sick{} is very easy to use, well-tested, parallelised, and freely available online through GitHub under the MIT license. 
\end{abstract}

\section{Introduction}
Most of our understanding of astrophysics has been interpreted from spectra. Given how informative spectroscopic data is to our understanding of astrophysics, it is not surprising that the last decade has seen a substantial increase of publicly accessible spectral data. Large scale surveys have driven this trend, each releasing in excess of hundreds of thousands \citep[e.g.,][]{wigglez,boss,segue,rave,gaia-eso} of spectra. Millions more spectra are expected in the coming years \citep[e.g.,][]{lamost,galah}.

The spectra are obtained from different astrophysical sources to meet specific scientific objectives, and the data vary in wavelength coverage, resolution, and noise distributions. For these reasons many collaborations expend significant resources to produce bespoke analysis software. Unfortunately this impedes scientific reproducibility, as many codes still remain closed-source more than a decade after the original article was published. Any comprehensive literature comparison subsequently becomes impossible, as systematics can be difficult to accurately characterise without in-depth knowledge of the methods or access to the software.

Broadly speaking there are three types of methods employed for spectral analysis: measuring the strengths of spectral features, pure data-generating models or template-matching methods. Approaches that measure spectral features are inexpensive, but regularly encounter problems with blended (often hidden) lines or continuum placement. As such, some subjective interaction, tuning, or ad-hoc `calibration' is almost always required. Data-generating methods compute model spectra at run-time, and whilst accurate, can be prohibitively expensive for large samples. For template-matching methods, synthetic spectra are only generated once to produce a grid of fluxes or intensities for a subset of permutations of astrophysical quantities. Although there are differences between these methods, the preparatory steps are usually the same. Spectra are placed at rest-frame by calculating line-of-sight velocities, typically by cross-correlation, before being continuum-normalised to flux intensities.

Credible uncertainties can be difficult to discern from these approaches. The implied assumptions of a $\chi^2$ distribution are often inapplicable, and uncertainties in the doppler-shift, smoothing, and normalisation steps are almost always ignored. These effects result in ill-characterised uncertainties in astrophysical parameters. Alternatively the uncertainties are frequently assumed to be approximately the same for all objects.  This is an incorrect approach: there are few, if any, examples of homoscedastic datasets in astrophysics. The noise properties of each spectrum \textit{are} different, and the parameter uncertainties (random and systematic) will differ for every object. Consequently the uncertainties in astrophysical parameters by template-matching methods are generally found to either be incorrectly assumed, under-estimated, or at least ill-characterised.

In addition to affecting the uncertainties, the effects of redshift, continuum normalisation and smoothing \textit{will} systematically bias the reported maximum-likelihood parameters. Experienced spectroscopists will frequently differ in their decision of continuum placement, even in the most straightforward cases (e.g., metal-poor stars). For example, there are a number of controversial examples within the literature where the subjective (human) decision of continuum placement have significantly altered the scientific conclusions \citep[e.g., see][where this issue is discussed in great detail]{kerzendorf}. The implications of these phenomena \textit{must} be considered if we are to understand subtle astrophysical processes. Spectroscopy requires objectivity: one should endeavour to incorporate these phenomena as free parameters into a generative model and infer them simultaneously with the astrophysical parameters.

In this \article{} I present \sick{}: a well-tested, MIT-licensed probabilistic software package for inference from spectroscopic data. \sick{} employs an approximation to data-generating models. Instead of attempting to model all kinds of expensive astrophysical processes (e.g., supernova explosions, stars, and any other interesting astrophysical processes) at run-time, it performs efficient multi-dimensional interpolation of model spectra from pre-computed grids. This approach is suitable for a plethora of different astrophysical processes, allowing any user to easily specify a model using existing published spectral grids, and infer astrophysical properties from their data. Aspects of the probabilistic model are described in Section \ref{sec:model}. A number of examples are included in Sections \ref{sec:inference-test} and \ref{sec:examples}, where I present accurate and precise objective photospheric measurements using noisy spectra. I conclude in Section \ref{sec:conclusions} with references to the online documentation and applicability of the software.

\section{The Generative Model}
\label{sec:model}

The primary advantage of \sick{} is the simultaneous inference of both astrophysical and other pertinent parameters that can alter the spectra. The generative model described here is agnostic as to \textit{what} the astrophysical parameters actually describe. Typical examples might be properties of supernova (e.g., explosion energies and luminosities), galaxy characteristics from integrated light, or mean plasma properties of a stellar photosphere. There are a plethora of spectral libraries (observed and synthetic) published for these types of applications \citep[e.g.,][]{snid,pegase,phoenix,pollux}. All of these models are fully-\sick{} compatible. However given the research background of the author, I will introduce the probabilistic model by focussing on stellar photospheric inferences.

The first step of the (approximate) generative model is to interpolate a model spectrum. \sick{} employs the Quickhull algorithm \citep{quickhull} to linearly interpolate in multiple dimensions. Irregularly-spaced grids are perfectly acceptable; Quickhull does not require the grid to be rectangular or regularly-spaced points. Large grids of model spectra are cached for computational efficiency, allowing for the total size of the model grid to far exceed the available random access memory.\footnote{In practice this is performed by efficient memory-mapping.} I introduce an interpolation function $S(\bm{\psi})$ that produces a model spectrum with $J$ points $\{\lambda_j,I_j\}_{0}^{J}$ at any evaluable point $\bm{\psi} \equiv (T_{\rm eff},\log{g},{\rm [M/H]},[\alpha/{\rm Fe}])$ within the grid boundaries.

\begin{equation}
S(\bm{\psi}) \rightarrow \{\lambda_j,I_j\}
\end{equation}

The model spectrum $S(\bm{\psi})$ should always be of higher spectral resolution than the data. Therefore it is necessary to smooth (or broaden) $\{\lambda_j,I_j\}$ with a Gaussian filter of standard deviation $\sigma_{s}$ such that it matches the spectral resolution of the data. For some applications it is tempting to perform this step \textit{a priori} from the instrument spectral resolution. However it is not necessarily true that the quoted instrumental resolution will perfectly match the data. In truth the data have a \textit{distribution} of spectral resolutions. Even if the spectral resolution is accurately known with low variance, fixing this value will result in under-estimated uncertainties $\sigma(\bm{\psi})$ and likely bias the posterior distribution distributions for $\bm{\psi}$. Employing the free parameter $\sigma_{s}$ allows for the identification of additional instrumental or atmospheric broadening. For stellar applications, the free parameter $\sigma_{s}$ also aids in the identification of unresolved binary companions.

The doppler shift of the source is treated by transforming the model spectrum (in uniformly sampled $\log\lambda$ space). \sick{} solves for redshift $z$ (Equation \ref{eq:redshift}), and returns posterior distributions in the user's preferred units (e.g., km s$^{-1}$).

\begin{equation}
\label{eq:redshift}
\lambda_{s} = \lambda_{j}(1 + z)
\end{equation}

After the model spectrum is smoothed and doppler shifted, model intensities are required at the wavelengths $\lambda_i$ of the observed pixels. The complete function to yield model intensities $M_i$ at wavelengths $\lambda_i$ is therefore a function of the parameters of interest $\bm{\psi}$, redshift $z$, smoothing kernel $\sigma_{s}$, and the wavelengths of the observed pixels $\lambda_i$:

\begin{equation}
 M_{i} \equiv S\left(\bm{\psi},z,\sigma_{s},\lambda_i\right)
\end{equation}

In principle there is no reason to believe that the doppler-shifted, smoothed model intensities $M_i$ should match the data at all. The counts and shape of observed fluxes are a function of source magnitude, exposure time, instrument sensitivities, atmospheric conditions, and a host of unaddressed effects. In contrast the model spectra are calculated either as intensities (e.g., $I_j$) or fluxes calibrated to an empirical system. Even for the true values of $\bm{\psi}$, a function is required to normalise\footnote{The normalisation process is frequently abused by stellar spectroscopists in the literature. Wherever possible, data should not be transformed. One should seek to fit a model to the data, not the other way around.} the model to the data. As such I transform the \textit{model} intensities by some function $C$ to fit the data. Although the function $C$ incorporates a number of effects (e.g., source blackbody temperature, dust, instrument sensitivities), they are phenomena that typically cannot be separated without additional information, and here I only care about their combined effect. I only wish to ensure that the overall shape of the data are accounted for, which is usually achievable with a low-order polynomial

\begin{equation}
C_i = c_{0}\lambda_{i}^{m-1} + c_{1}\lambda_{i}^{m-2} + \dots + c_{m}
\end{equation}

\noindent{}with $m$ free coefficients $c_{0\dots{}m}$, where $m$ is specified by the user. Other continuum functions are available in \sick{}, and the user is encouraged to experiment for a specific problem. This allows us to express the \textit{expected flux} $E_i$ at a given observed pixel with wavelength $\lambda_i$ as:

\begin{equation}
E_i = S_{i}\times{}C_i
\end{equation}

In many astrophysical cases the flux uncertainties $\sigma_{i}$ for a given pixel are not well-characterised. This is often due to unpropagated uncertainties during data reduction. Since the observed flux counts are expected to be Poisson-distributed, in many astrophysical scenarios the pixel flux uncertainty can be made estimated as $\sigma_i \sim 1/\sqrt{F_{i}}$. Knowing that this approximation may underestimate the noise, I include an additional parameter to account for the possibility that the variance in all pixels is underestimated by some fractional amount $f$ (Equation \ref{eq:noise-model}). While this crude noise model is probably unrepresentative of the data in a large number of cases, including \textit{any} noise model is preferable to none.

\begin{equation}
s_{i}^2 = \sigma_{i}^2 + f^{2}E_{i}^2
\label{eq:noise-model}
\end{equation}

We now have a \textit{generative model} for the data. The frequency (or probability) distribution ${p\left(F_i|\lambda_i,\sigma_i,\bm{\psi},z,\sigma_s,\{c\}_{0}^{m}\right)}$ for $F_i$ is given by

\begin{equation}
p\left(F_i|\lambda_i,\sigma_{i},\bm{\psi},z,\sigma_{s},\{c\}_{0}^{m},f\right) = 
 \frac{1}{\sqrt{2\pi{}s_{i}^2}}\exp{\left(-\frac{\left[F_i - E_i\right]^2}{2s_{i}^2}\right)}
 \label{eq:p_model}
\end{equation}

\noindent{}and with the implied assumption that the data are independently drawn, the likelihood $\mathcal{L}$ is calculable by the product of individual probabilities:

\begin{equation}
\mathcal{L} = \prod_{i=1}^{N}\,p\left(F_i|\lambda_i,\sigma_{i},\bm{\psi},z,\sigma_{s},\{c\}_{0}^{m},f\right)
\end{equation}

From the original astrophysical parameters $\bm{\psi}$ that I care about, I now have an additional $3 + m$ parameters to consider. The situation seemingly becomes more complex when separately observed channels are considered. Many spectrographs provide small portions of spectra (a channel, also colloquially expressed as beams or apertures) separated by some large gap. The data for each channel are usually captured by different CCDs and are therefore reduced separately.

The instrumental parameters $\sigma_{s}$, $\{c_k\}_{k=0}^{m}$, and $f$ are likely to be different for each channel. Although redshift $z$ is an astrophysical effect and should not differ between channels, this may not be true if each channel has been wavelength-calibrated differently. It is therefore prudent to introduce separate parameters $z$, $\sigma_{s}$, $\{c_k\}_{k=0}^{m}$, $f$ for \textit{each} of the $N_{chan}$ observed channels. This scales the total dimensionality of the model as $3\times{}N_{chan} + \sum_{k=0}^{N_{chan}}m_{k}$ in addition to $\bm{\psi}$. The inclusion of all of these parameters is not mandatory: each \sick{} model can be adjusted to include or ignore any combination of phenomena.

Finally, I consider the handling of outliers in the data. These may be in the form of cosmic ray spikes, improper calibration of the data, telluric features, or simply poorly modelled spectral regions. Treatment of these artefacts is achieved using a Gaussian mixture model: a combination of two models. In the mixture model, the data are fit by the sum of amplitudes ($1 - P_o$ and $P_o$, respectively) of two distributions: the expected fluxes $E_i$, and a normal distribution centered along the continuum function $C_i$ with variance $s_{i}^2 + V_{o}$. This requires the inclusion of two (final) additional parameters: $P_o$ and $V_o$. The prior distribution function $p\left(V_o\right)$ requires $V_{o}$ to always be positive (Equation \ref{eq:default_priors}), and as such the outlier distribution will \textit{always} have a larger variance. Distributions of smaller variance are more informative, so conceptually a fit to the expected fluxes $E_{i}$ is generally preferred wherever possible. For brevity I define $\bm{\kappa} \equiv (\bm{\psi},\{z,\sigma_s,\{c_k\}_{k=0}^{m},f\}_{0}^{N_{c}})$, and the likelihood for the mixture model is given by

 \begin{equation}
\mathcal{L} = \prod_{i=1}^{N}\,\left[\left(1 - P_{o}\right)\times{}p_{model}\left(F_i|\lambda_i,\sigma_{i},\bm{\kappa}\right) + P_{o}\times{}p_{outlier}\left(F_i|\lambda_i,\sigma_i,\bm{\kappa},V_{o},P_o\right)\right]
\end{equation}

\noindent{}where $p_{model}$ refers to $p$ in Equation \ref{eq:p_model} and

\begin{equation}
p_{outlier}\left(F_i|\lambda_i,\sigma_i,\bm{\kappa},V_{o},P_o\right) = \frac{1}{\sqrt{2\pi\left(s_{i}^2 + V_{o}^2\right)}} \exp\left(-\frac{[F_i - C_i]^2}{2\left[s_{i}^2 + V_{o}^2\right]}\right)
\end{equation}

\noindent{}such that the likelihood $\mathcal{L}$ becomes:

\begin{equation}
\mathcal{L} = \prod_{i=1}^{N} \left[ \frac{1-P_o}{\sqrt{2\pi{}s_{i}^2}}\,\exp\,\left(-\frac{[F_i - E_i]^2}{2s_{i}^{2}}\right) + \frac{P_o}{\sqrt{2\pi\left[s_{i}^2 + V_o\right]}}\,\exp\,\left(-\frac{[F_i - C_i]^2}{2\left[s_{i}^{2} + V_o\right]}\right)\right]
\label{eq:full_likelihood}
\end{equation}

I define the full parameter space with $\bm{\theta} \equiv \left(\bm{\psi},\{z,\sigma_s,\{c_{b_k}\}_{k=0}^{m},f\}_{b=0}^{N_{c}},V_o,P_o\right)$. From Bayes theorem the posterior probability distribution for $\bm{\theta}$ (up to a constant) is given by
\begin{eqnarray}
\mathcal{P} & \propto & likelihood \times prior \nonumber \\
p(\bm{\theta}|\{F_i\}_{i=1}^{N}) & \propto & p(\{F_i\}_{i=1}^{N}|\bm{\theta})\,\times\,p(\bm{\theta})
\label{eq:probability}
\end{eqnarray}

\noindent{}where $p(\bm{\theta}|\{F_i\}_{i=1}^{N})$ is the probability $\mathcal{P}$ of $\bm{\theta}$ given the data (and given the model: e.g., see Section \ref{sec:sun}), $p(\{F_i\}_{i=1}^{N}|\bm{\theta})$ is our previously defined likelihood function $\mathcal{L}$ (Equation \ref{eq:full_likelihood}), and $p(\bm{\theta})$ is the prior probability distribution. Priors are discussed in more detail in Section \ref{sec:mcmc}.

% Log probability, expand function
%\begin{eqnarray}
%\log({\mathcal{P}}) & \propto & \log{likelihood} + \log{prior} \nonumber \\
%\log(\mathcal{P}) & \propto & \prod_{i=1}^{N} \left[ \frac{1-P_b}{\sqrt{2\pi\sigma_{i}^2}}\,\exp\,\left(-\frac{[F_i - E_i]^2}{2\sigma_{i}^{2}}\right) + \frac{P_b}{\sqrt{2\pi\left[\sigma_i^2 + V_o\right]}}\,\exp\,\left(-\frac{[F_i - C_i]^2}{2\left[\sigma_i^{2} + V_o\right]}\right)\right] \nonumber \\
%& & \dots +  \ln(p(\bm{\theta}))
%\end{eqnarray}

With all of the combined effects there are 16 free parameters including $\bm{\psi}$, our astrophysical quantities of interest. For the examples presented in Section \ref{sec:examples}, this requires interpolation between 92 billion pixels in four dimensions. While the description might appear daunting, the problem is tractable, numerically efficient, and easy to configure. By default \sick{} numerically solves this problem in three sequential steps, which are described in the following sections: scattering, optimisation, and Monte-Carlo Markov Chain (MCMC) sampling.

\subsection{Initial Scattering}
\label{sec:scattering}

We seek to maximise $\mathcal{P}$ (or in practice, $\log\mathcal{P}$) and calculate the posterior probability distributions for $\bm{\theta}$ given the data. In the first step I calculate the log-probability $\log{(\mathcal{P})}$ for $N_{sample}$ randomly drawn points from all over the astrophysical parameter space $\bm{\psi}$. Since \sick{} allows for arbitrarily large parameter spaces in $N_{D}$ dimensions, initially sampling the parameter space $N_{sample}$ times provides a coarse landscape of probability. If the probability distribution function were smoothly distributed across all $\bm{\psi}$, or the optimisation step was sufficiently robust for all potential applications, then in principle \textit{any} single point would be an adequate starting guess. However there is no way of knowing \textit{a priori} how smooth the posterior distribution will be for a given problem.

The $N_{samples}$ points are uniformly drawn in $\bm{\psi}$ by default, unless priors are specified (see Section \ref{sec:mcmc}). A model spectrum $S(\bm{\psi})$ is interpolated for each point, which is used to estimate the redshift and continuum parameters. After smoothing the model fluxes $I_j$ by $\sigma_s$ (as drawn from an explicit prior, or estimated by $\left|\mathcal{N}\left(0, 1\right)\right|$) the smoothed spectrum is cross-correlated with the data to yield a redshift estimate. Continuum parameters $\{c_k\}_{k=0}^{m}$ are similarly estimated by fitting a function (a polynomial in this case) to the data divided by the smoothed flux at each $\lambda_i$. Estimating continuum and redshift provides us with a reasonable guess of the probability for any randomly scattered point $\bm{\psi_i}$. If outlier modelling is included in the model then $P_o$ is distributed as $\mathcal{U}\left(0, 1\right)$ by default, and $V_o$ is estimated by $\mathcal{N}\left(\widetilde{F_i}, \frac{1}{2}\widetilde{F_i}\right)$. This procedure is \textit{only} employed for the initial random scattering stage: it is \textit{not} applicable for the MCMC phase.

\subsection{Optimisation}
\label{sec:optimise}

After the random scattering step, the most probable $\bm{\theta}$ point is used as an initial guess for numerical optimisation. A number of suitable minimisation algorithms are available in \sick{} through the SciPy \citep{scipy} optimization module. The Nelder-Mead algorithm \citep{nelder-mead} is employed by default, and the full suite of optimisation algorithms available is listed in the online documentation. Unlike other optimisation techniques, the Nelder-Mead algorithm does not approximate first- or second-order derivatives, and can consequently be less efficient than other approaches. However it is robust in high-dimensional space, even in the presence of substantial noise, and has been successfully employed in a wide range of engineering and scientific problems. The reader is encouraged to experiment with other optimisation algorithms if the Nelder-Mead algorithm proves unsuitable or takes an untenable amount of time. Since these optimisation algorithms are minimisation techniques and I seek to maximise the log-probability $\log{\left(\mathcal{P}\right)}$, I numerically optimise the parameters $\bm{\theta}$ by minimising the negative log-probability $-\log{\left(\mathcal{P}\right)}$.

\subsection{Monte-Carlo Markov Chain Sampling}
\label{sec:mcmc}

The random scattering and numerical optimisation steps efficiently provide an accurate estimate of the optimal parameters $\bm{\theta}$. Once the optimisation step is complete, \sick{} employs the affine-invariant ensemble sampler proposed by \citet{goodman;weare}, and implemented by \citet{emcee}. The Metropolis-Hastings MCMC algorithm is employed by default. \sick{} allows for the model settings to be specified in a human-readable \textsc{yaml}- or \textsc{json}-formatted configuration file\footnote{The reader is referred to the online documentation for an example.}, where the number of Goodman \& Weare walkers can be specified, as well as the number of samples to perform. When the optimisation step is used, the initial points are taken from a small multi-dimensional ball around the optimised parameters $\bm{\theta_{opt}}$. If the optimisation step is not performed, initial values for the MCMC walkers are drawn in the same way as described for the scattering process.

Priors represent our initial knowledge about a particular parameter before looking at the data, and are necessary for any Bayesian analysis. A number of different prior distributions can be specified by the user in the \sick{} model configuration file. When no prior is explicitly specified, the following uninformative prior distributions are assumed (for all channels, where appropriate):

\begin{eqnarray}
p\left(\bm{\psi}_{dim}\right) &=& \mathcal{U}\left(\min\left[\bm{\psi}_{dim}\right], \max\left[\bm{\psi}_{dim}\right]\right) \\
p\left(P_o\right) &=& \mathcal{U}\left(0, 1\right) \\
p\left(\log{f}\right) &=& \mathcal{U}\left(-10, 1\right) \\
p\left(V_o,\sigma_s\right) &=& \left\{
\begin{array}{c l}      
    1\,, &\mbox{for values greater than zero}\\
    0\,, &\mbox{otherwise}
\end{array}\right. \\
p\left(z,\{c_k\}_{k=0}^{m}\right) &=& 1 \\
\label{eq:default_priors}
\end{eqnarray}

A consequence of allowing irregular model grids is that occasionally a model spectrum cannot be interpolated for some values of $\bm{\psi}$, even if they fall within $\left(\min\left[\bm{\psi}\right], \max\left[\bm{\psi}\right]\right)$. In these cases $p\left(\bm{\psi}\right) = 0$ and thus $\log\left(\mathcal{P}\right) = -\infty$.

Quantitatively ensuring numerical convergence for MCMC analyses remains an unsolved problem. There are a number of excellent resources on MCMC sampling which outline this issue in greater detail. The mean acceptance fraction, auto-correlation times, and values of the parameter chains themselves all provide reasonable proxy indicators of convergence. The user is encouraged to inspect these metrics and devise some heuristic for when convergence has unambiguously been achieved. The final state of every MCMC analysis is automatically saved by \sick{}. This allows users to resume their analysis from the most recent state if they believe the system has not converged.

\subsection{Self-consistent inference test}
\label{sec:inference-test}

As an initial test of the probabilistic framework I have interpolated a noise-free synthetic spectrum which will act as a faux observation. A single channel from the AMBRE synthetic library \citep{ambre} has been employed. The spectra were calculated for the Gaia-ESO Survey \citep{gaia-eso} and range from 475\,nm to 685\,nm. I have applied a number of transformations to the interpolated spectrum: a second-order polynomial enters multiplicatively to represent the continuum, the spectrum is redshifted, and fluxes are convolved to a spectral resolution of $\mathcal{R} \sim 10000$. The faux data are resampled to a uniform spacing of $\sim{}0.8$\,\AA{}, effectively discarding $99.5\%$ of the original pixels, which were sampled at $\sim4\times10^{-4}$\,nm. White noise has been added to replicate a S/N ratio of $\sim 7$ pixel$^{-1}$. I also assumed that the observed variance is systematically underestimated by 10 per cent: conceptually the faux data is $\sim{}$10 per cent noisier than what an observer would estimate from the flux counts.

% The AMBRE synthetic spectra have been calculated using 1D MARCS model atmospheres and synthesised using Turbospectrum \citep{turbospectrum} under the assumption of local thermal equilibrium (LTE). There are 28,344 permutations of effective temperature $T_{\rm eff}$, surface gravity $\log{g}$, overall metallicity $[{\rm M/H}]$ and mean $\alpha$-element enhancement $[\alpha/{\rm Fe}]$. The typical step size is 250 K in temperature, 0.25 dex in $\log{g}$, 0.25 dex in [Fe/H] and 0.10 dex in [$\alpha$/Fe]. Each grid point has synthesised continuum-normalised intensities between [0, 1]  for two spectral channels: one from 475\,nm to 685\,nm, and a redder channel covering 845\,nm to 895\,nm. Both channels are sampled at approximately $\sim4\times10^{-4}$\,nm, yielding more than $92\times10^9$ pixels.

Default priors and configurations were employed. 200 walkers were used to explore the parameter space for 1250 steps as burn-in, and another 250 steps were used to sample the posterior. The chains for each parameter are shown in Figure \ref{fig:chains}. At least by eye, the simulation appears to have converged after just 300 steps. The marginalised posterior distributions are shown in Figure \ref{fig:corner-inference}, with the accepted truth values marked. The true and inferred values for all parameters are tabulated in Table \ref{tab:inference-test}.

This test demonstrates that accurate and precise inference of stellar parameters can be made even in the presence of substantial noise. The true values of all parameters are recovered within the uncertainties, and the largest (relative) uncertainties are observed for continuum coefficients $b_1$ and $b_2$. Figure \ref{fig:corner-inference} shows there are strong covariances between these parameters. This is unsurprising given the shape of the spectrum (Figure \ref{fig:spectrum-inference}), which could reasonably be fit with a linear continuum, even though the continuum was truly modelled as a third order polynomial. The logarithm of additional fractional variance $\ln{f_{blue}}$ is precisely constrained, suggesting the noise model is excellently matched by the data. Given the simplicity of the noise model, the precision of inference on $\ln{f_{blue}}$ is unlikely to be recovered for real data.

\begin{figure}
\label{fig:chains}
\includegraphics[height=\textheight]{chains.pdf}
\caption{Points sampled by the 200 walkers at each step during the self-consistent inference test. The true values are marked in blue. The first 1250 steps are discarded as the burn-in period.}
\end{figure}

\begin{figure}
\label{fig:corner-inference}
\includegraphics[width=\textwidth]{corner.pdf}
\caption{Marginalised posterior distributions for all parameters $\bm{\theta}$ from a faux observation with spectral resolution $\mathcal{R} \sim 10,000$ and S/N ratio $\sim{}7$\,pixel$^{-1}$. True values are marked in blue. This figure demonstrates that precise inference of stellar parameters can be made with high-resolution spectra, even in the presence of substantial noise.}
\end{figure}

\begin{table}
\caption{True and inferred model parameters for the self-consistent inference test}
\label{tab:inference-test}
\begin{tabular}{llrr}
\hline
\hline
Parameter & Description & Truth & Inferred \\
\hline
$T_{\rm eff}$ & Effective photospheric temperature (K) & 5454 & 5483$_{-137}^{+125}$ \\
$\log{}g$ & Surface gravity & 4.124 & 4.084$_{-0.274}^{+0.262}$ \\
${\rm [Fe/H]}$	& Metallicity & --0.514 & $-0.512_{-0.160}^{+0.156}$ \\
$[\alpha/{\rm Fe}]$ & $\alpha$-element abundance & $+0.02$ & $+0.05_{-0.12}^{+0.13}$ \\
$v$		& Velocity (km s$^{-1}$) & 13.0 & $9.5_{-5.5}^{+5.3}$ \\
$\sigma_{blue}$ & Gaussian smoothing sigma (\AA{}) & 0.581 & 0.518$_{-0.34}^{+0.39}$ \\
$\ln{f_{blue}}$ & Logarithm of fractionally underestimated variance & $-2.30$ & $-2.28_{-0.02}^{+0.02}$ \\
$b_{0}$ & Continuum polynomial coefficient ($\times10^{-3}$) & 1.23 & $1.26_{-0.32}^{+0.32}$ \\
$b_{1}$ & Continuum polynomial coefficient & $-0.593$ & $-1.001_{-3.634}^{+3.681}$ \\
$b_{2}$ & Continuum polynomial coefficient & $-0.569$ & $894_{-10335}^{+10367}$ \\
\hline
\end{tabular}
\end{table}

\begin{figure}
\label{fig:spectrum-inference}
\includegraphics[width=\textwidth]{spectrum.pdf}
\caption{A faux observed spectrum (black) of $\mathcal{R} \sim 10000$ and S/N ratio of $\sim7$ with an underestimated variance by $\sim$10\%. The maximum likelihood model spectrum is shown in red. }
\end{figure}

\section{Utility}
\label{sec:examples}
Here I include two realistic applications to demonstrate the capabilities and ease-of-use of the software.

\subsection{Sol}
\label{sec:sun}
A high-resolution ($\mathcal{R} \sim$ 20000), high S/N ($\sim{}$150 pixel$^{-1}$) twilight spectrum was obtained from the GIRAFFE/FLAMES archive\footnote{https://www.eso.org/observing/dfo/quality/GIRAFFE/pipeline/solar.html}. The H875.7 setup was chosen for this example, yielding spectra from 848.2\,nm to 899.2\,nm. Data points redder than 874\,nm have been discarded to more closely match the Gaia RVS wavelength coverage, and demonstrate the recoverability of stellar parameters on real data.

The AMBRE \citep{ambre} synthetic grid was also employed for this example. The continuum was represented with a third order polynomial, and no redshift or outlier modelling was included. The central $\pm$0.1 nm cores of the Ca II near-infrared triplet lines were masked (specifically rest wavelengths 849.7-849.9 nm, 854.1-854.3, and 866.1-866.3 nm), as the cores are strongly affected by non-LTE effects that are not treated in the 1D LTE model atmospheres used to generate the AMBRE spectra. Pixels between rest wavelengths 850.10\,nm to 850.37\,nm were also masked as the models did not seem to accurately reproduce the data in this region.

\begin{figure}
\label{fig:solar}
\includegraphics[width=\textwidth]{solar.pdf}
\caption{Stellar parameter ($\bm{\psi}$: $T_{\rm eff}$, $\log{g}$, [Fe/H], [$\alpha$/Fe])  posterior distributions for the GIRAFFE/FLAMES twilight spectrum. Nuisance parameters are not shown. Maximum likelihood values for each parameter are drawn in red, and the 16\%, 50\% and 84\% quantiles are shown as dashed lines. The posterior distributions are precise, and close to the accepted values (not shown) of $T_{\rm eff} = 5777$\,K, $\log{g} = 4.445$, slightly discrepant from [$\alpha$/Fe] = 0, and significantly deviant from [Fe/H] = 0. This test illustrates the precision achievable using \sick, but highlights the need for accurate models.}
\end{figure}

For this example 1000 draws were made during the random scattering stage, and optimisation was performed using the default options. One hundred walkers explored the parameter space for 1000 steps, before another 200 steps were used to sample the posterior. The posterior distributions for stellar parameters is shown in Figure \ref{fig:solar}. The currently accepted stellar parameters for the Sun are $T_{\rm eff} = 5777$\,K, $\log{g} = 4.445$, [Fe/H] = 0, and [$\alpha$/Fe] = 0. The inferred parameters agree excellently with the accepted values for effective temperature and surface gravity: $T_{{\rm eff},inferred} = 5770^{+20}_{-14}$, and $\log{g}_{inferred} = 4.40^{+0.06}_{-0.05}$. The agreement is less pleasing for $\alpha$-element enhancement, but less so for mean metallicity: [$\alpha$/Fe]$_{inferred} = -0.03^{+0.01}_{-0.01}$, and [Fe/H]$_{inferred} = -0.07^{+0.01}_{-0.01}$.

While our inferences are precise, the offset in mean metallicity suggest the models may be slightly inaccurate in this wavelength regime or that there are unaccounted telluric features in the data. This test highlights the need to verify the accuracy of the input models. A useful extension of this project would be to simultaneously include telluric spectra, as well as pre-computed astereoseismic models, such that spectroscopy and time-series photometry could simultaneously constrain stellar parameters. Asteroseismology can accurately and precisely constrain surface gravity, but requires an input effective temperature. Conversely, surface gravity is the least constrained parameter by spectroscopy, but can precisely constrain effective temperature is the only method available to measure chemical abundances. Since \sick{} allows for any dispersion unit equivalency (e.g., frequency, wavelength, etc), spectroscopic and astereoseismic models can be trivially coupled into a generative framework for unprecedented accuracy and precision in stellar spectroscopy.

\subsection{Globular Clusters}
Globular cluster stars are excellent candidates for testing \sick{} over a large range of stellar parameters. Here I consider a realistic scenario of low-resolution, noisy spectra of red giant and asymptotic branch cluster candidates observed using the AAOmega spectrograph on the Anglo-Australian Telescope in Coonabarabran, Australia. The 580V and 1700D gratings were employed, yielding one channel between 370\,nm to 580\,nm with a spectral resolution $\mathcal{R} \sim 1300$, and another from 820\,nm to 890\,nm with spectral resolution $\mathcal{R} \sim 8000$.

Although these data are generally considered to be of low spectral resolution, I show that astrophysical parameters can be accurately inferred with high precision. The AMBRE synthetic grid has also been employed for this example. Each observed channel was treated identically: individual doppler shifts were permitted in each channel, and the continuum was represented as a second-order polynomial. Outlier modelling was included.
The parameter space was randomly sampled 1000 times before numerical optimisation was performed using the Nelder-Mead algorithm. 200 walkers explored the parameter space for a burn-in period of 450 steps, before sampling the posterior another 50 times. The auto-correlation times, sampled chain values, and mean acceptance fractions demonstrated that convergence was unambiguously achieved for all analysed sources.

Radial velocities were used to identify and discard non-cluster members. The metallicity distribution of the remaining cluster stars are shown in Figure \ref{fig:clusters}, where a comparison to the literature is made. The \citet{harris} catalogue (accessed June 2014) has been used as a literature reference for all globular clusters. Note that while there are measurable intrinsic metallicity spreads for these clusters, I have not adopted these as x-axis uncertainties.

\begin{figure}
\label{fig:clusters}
\includegraphics[width=\textwidth]{clusters.pdf}
\caption{Metallicities of cluster stars inferred from noisy, low-resolution spectra obtained with the AAOmega spectrograph. The agreement with the literature values (see text) is very reasonable.}
\end{figure}

Even ignoring uncertainties in literature values, the agreement between is exceptional for noisy, low-resolution spectra. No systematic trend is present and a negligible mean metallicity difference of $-0.001$ dex (accuracy of 0.2\%) is observed over three orders of magnitude. The agreement would be remarkable for high-resolution spectra with exquisite S/N ratios. I emphasise that these results were obtained without performing any `calibration' or ad-hoc posterior corrections, which are usually entirely empirical and not motivated by an astrophysical understanding.

Individual uncertainties in stellar parameters are also extremely reasonable, particularly given the low spectral resolution and S/N of the data. Uncertainties of $\sim$75 K in $T_{\rm eff}$, $\sim0.2$ in $\log{g}$, $\sim$0.1 dex in [Fe/H] and $\sim0.05$ dex in [$\alpha$/Fe] are obtained in the presence of substantial noise (e.g., $S/N \sim 15$). The most uncertain results were obtained for NGC 7089, where the lowest S/N ratios exist and only 6 cluster members were observed.

\section{Conclusion}
\label{sec:conclusions}

A probabilistic code has been introduced to approximately forward model spectroscopic data, and thereby allow for simultaneous inference of astrophysical and nuisance parameters. The generative model approach described here has a number of advantages over previously published techniques. Preparatory and subjective decisions (e.g., redshift and placement of continuum) are objectively treated within a scalar-justified mathematical model. This allows users to credibly characterise the parameter posterior distributions and understand how they affect astrophysical measurements. Until now most (if not all) published techniques treat these processes separately, thereby systematically biasing and generally under-estimating the parameter uncertainties.

With applications to both high- and low-resolution stellar spectra, I have demonstrated that the simultaneous incorporation of these effects leads to remarkable improvements in both accuracy and precision. While the examples presented here have focussed on stellar spectra, the code is ambivalent about \textit{what} the astrophysical parameters describe: the framework can be easily used for any kind of quantifiable astrophysical process. The code is MIT-licensed (freely distributable, open source) and has an extensive automatic testing suite. Complete documentation is available online\footnote{github.com/andycasey/sick/wiki}, which includes a number of additional examples and tutorials. In the spirit of full scientific reproducibility, the online documentation is complemented with the files necessary to reproduce all of the examples presented in this \article{}. The source code is distributed using \textsc{git}\footnote{git-scm.com/downloads}, and hosted online at GitHub\footnote{github.com/andycasey/sick}.

Great care has been spent to ensure the code is easy to use, and allow users to obtain precise inference with little effort. I strongly encourage the use of this software for existing and future spectral datasets. With the sheer volume and high-quality of spectral data available, astronomers must begin to adopt objective, generative models for their data. Subtle astrophysical processes can only be discovered and understood with the proper characterisation of uncertainties afforded by generative models.

\acknowledgements
I am pleased to thank Louise Howes for extensive testing of this code, Martin Asplund for providing commentary, as well as Sergey Koposov and Jarryd Page for constructive discussions on this work. This research has been funded by European Research Council grant 320360: The Gaia-ESO Milky Way Survey.

% ADS?
% Coveralls?
% GitHub?
% Triangle.py?

\bibliographystyle{apj}
\bibliography{biblio}
\end{document}